RILS-ROLS: Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-09-17 doc/paper.tex
ILS: explained

Koza~\cite{koza1994genetic} introduced the problem of SR as a specific application of genetic programming. GP deals with the object called programs, which need to be optimized. In particular, the programs are represented by syntax trees consisting of functions/operations over input features and constants; as an example of a function represented by a syntax tree, see Figure~\ref{fig:syntax-tree-example} In essence, syntax trees are elements of the search space of SR. That is, each sample model $\tilde{f}$ may be seen as a point in the search space, represented by respective syntax tree. How accurate is this mapping, we compute in the basis of the historical data $D$ and the chosen error   measure (such as $MSE$, or $R^2$). Interestingly, the search space of SR consists of its discrete and continuous part. More precisely, the model we are seeking for might be obtained as a composition of some functions from a (finite) set of mathematical functions. It is common to use the set of the following elementary mathematical functions in the experiments: $\sqrt{x}, x^2 $, $\sin$, $\cos$, $\log$, $\exp$, $\arcsin$, $\arccos$, and $a^x$. It is allowed to link functions will all four operators: $+$, $-$, $\times$, and $/$. After fixing functions that build models, the set of coefficients that  fits to this model must be provided. For the most physical models, these coefficients may receive any real value, that is they are continuous in its nature.

\fxnote{TODO: proceed...}

@@

   \subsection{Iterated local search}
   \fxnote{TODO: Marko}

>>>>

Koza~\cite{koza1994genetic} introduced the problem of SR as a specific application of genetic programming. GP deals with the object called programs, which need to be optimized. In particular, the programs are represented by syntax trees consisting of functions/operations over input features and constants; as an example of a function represented by a syntax tree, see Figure~\ref{fig:syntax-tree-example} In essence, syntax trees are elements of the search space of SR. That is, each sample model $\tilde{f}$ may be seen as a point in the search space, represented by respective syntax tree. How accurate is this mapping, we compute in the basis of the historical data $D$ and the chosen error   measure (such as $MSE$, or $R^2$). Interestingly, the search space of SR consists of its discrete and continuous part. More precisely, the model we are seeking for might be obtained as a composition of some functions from a (finite) set of mathematical functions. It is common to use the set of the following elementary mathematical functions in the experiments: $\sqrt{x}, x^2 $, $\sin$, $\cos$, $\log$, $\exp$, $\arcsin$, $\arccos$, and $a^x$. It is allowed to link functions will all four operators: $+$, $-$, $\times$, and $/$. After fixing functions that build models, the set of optimal coefficients that  fits to this model must be provided. For the most physical models, these coefficients may receive any real value, that is they are continuous in its nature.

\fxnote{TODO: needs for extension...}

@@

   \subsection{Iterated local search}
    Iterated local search (ILS)~\cite{lourencco2003iterated} is an efficient meta-heuristics, whose essential idea is given as follow. Iteratively generate a sequence of solutions generated by the (embedded) heuristic, such as local search or  randomized greedy heuristics,  leading
    towards better solutions than if one uses repeated random trials of that heuristic. This simple idea appeared in the literature by Baxter~\cite{baxter1981local} in early 1980s, and, since then, had been many time re-invented by many researches under different names; some of the following names were used: iterated descent search~\cite{baum1998iterated}, large-step
    Markov chains~\cite{martin1991large}, chained local optimization~\cite{martin1996combining}, or, in some cases,  combinations of these~\cite{applegate2003chained}.

   In the literate, one of the mostly applied ILS versions in the literature is pseudo-coded in Algorithm~\ref{alg:ils}. Note that we also used the core of this version in our \textsc{Risl-Rols} algorithm. \fxnote{Provjeriti da li je ovo tačno -- popraviti algoritam ako nije sa onim što je korišten...}

   \begin{algorithm}
  		\begin{algorithmic}[1] 
  			 \State \textbf{Input}: a COP problem; an LS procedure  
  	 		\State \textbf{Output}: an approximate (feasible) solution 
  	 		\State $s \gets$ Generate an initial solution
  	 		\State  $s \gets LS(s)$
   	 		\While{\emph{Stopping criteria is not met}}
   	    		 \State  $s' \gets$ Perturbation($s$)
   	    		 \State  $s' \gets$ LS($s'$)
   	    		 \State  $ s \gets$ AcceptanceCriterion($s, s'$)
   	 		\EndWhile
   		\end{algorithmic}
    	\caption{An ILS framework from the literature.}
    	   	     \label{alg:ils}
    \end{algorithm}

An initial solution may be generated randomly or by using a greedy heuristic, which is afterwards, improved by LS procedure. At each iteration, ILS applies three steps. First, the current incumbent solution is perturbed, that is mutated. The perturbed solution is then improved by LS procedure. The outcome is thereafter compared to the incumbent solution and possibly replaced for the new one, with accordance to the chosen acceptance criterion.
